---
title: "Olympic Games"
format: 
  dashboard:
    orientation: columns
    nav-buttons: [github]
    github: https://github.com/posit-conf-2024/olympicdash
logo: images/olympics-logo.svg
logo-alt: "Olympics logo with multicolored circles."
---

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
library(gt)
```

```{r}
#| label: load-data
#| message: false
olympics_full <- read_csv("../data/olympics.csv")
```

```{r}
#| label: prep-data
olympics <- olympics_full |>
  filter(!is.na(medal)) |>
  separate_wider_delim(cols = team, names = c("team", "suffix"), delim = "-", too_many = "merge", too_few = "align_start") |>
  select(-suffix) |>
  mutate(medal = fct_relevel(medal, "Bronze", "Silver", "Gold"))
```

# 🌞 Summer Olympics

```{r}
#| label: summer-prep-data
summer_olympics <- olympics |>
  filter(season == "Summer")
```

## Column - Medals by sport and year {width=65%}

### Row - Medals by sport {height=60%}

```{r}
#| label: summer-medals-by-sport
#| title: Medals by sport
#| fig-width: 10
#| fig-asp: 0.618
summer_olympics |>
  mutate(
    sport = fct_lump_n(sport, n = 15),
    sport = fct_infreq(sport),
    sport = fct_rev(sport),
    sport = fct_relevel(sport, "Other", after = 0)
  ) |>
  ggplot(aes(y = sport, fill = medal)) +
  geom_bar() +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```

### Row - Medals by year {height=40%}

::: {.card title="Medals by year"}
Due to World War II, no olympic games were held in 1940 and 1944.

```{r}
#| label: summer-medals-by-year
#| fig-width: 10
#| fig-asp: 0.3
summer_olympics |>
  count(year, medal) |>
  ggplot(aes(x = year, y = n, color = medal)) +
   geom_point(size = 0.8) +
  geom_line() +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_x_continuous(breaks = seq(1896, 2020, 8)) +
  labs(
    x = "Year",
    y = NULL,
    color = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = c(0.2, 0.8),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```
:::

## Column - Medals by country {width=35%}

### Row - Value boxes {height=30%}

```{r}
#| label: summer-calculate-most-medals
summer_most_golds <- summer_olympics |>
  filter(medal == "Gold") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1)

summer_most_silvers <- summer_olympics |>
  filter(medal == "Silver") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1)

summer_most_bronzes <- summer_olympics |>
  filter(medal == "Bronze") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1)
```

::: {.valuebox icon="award-fill" color="#d4af37"}
Most golds: 

`{r} summer_most_golds$n`

`{r} as.character(summer_most_golds$team)`
:::

::: {.valuebox icon="award-fill" color="#c0c0c0"}
Most silvers: 

`{r} summer_most_silvers$n`

`{r} as.character(summer_most_silvers$team)`
:::

::: {.valuebox icon="award-fill" color="#cd7f32"}
Most bronzes: 

`{r} summer_most_bronzes$n`

`{r} as.character(summer_most_bronzes$team)`
:::

### Row - Tabsets of tables {height=70% .tabset}

::: {.card title="Top 30 total medals"}
Teams sorted in descending order of total medals.

```{r}
#| label: summer-top-30-total-medals
summer_olympics |>
  count(team, medal) |>
  pivot_wider(names_from = medal, values_from = n, values_fill = 0) |>
  mutate(total = Bronze + Gold + Silver) |>
  arrange(desc(total), team) |>
  slice_head(n = 30) |>
  select(Team = team, Gold, Silver, Bronze) |>
  gt() |>
  cols_align(align = "left", columns = Team) |>
  data_color(
    method = "numeric",
    palette = "nord::aurora"
  )
```
:::

::: {.card title="Bottom 30 total medals"}
Teams sorted in ascending order of total medals.

```{r}
#| label: summer-bottom-30-total-medals
summer_olympics |>
  count(team, medal) |>
  pivot_wider(names_from = medal, values_from = n, values_fill = 0) |>
  mutate(total = Bronze + Gold + Silver) |>
  arrange(total, team) |>
  slice_head(n = 30) |>
  select(Team = team, Gold, Silver, Bronze) |>
  gt() |>
  cols_align(align = "left", columns = Team) |>
  data_color(
    method = "numeric",
    palette = "nord::frost"
  )
```
:::

# ❄️ Winter Olympics

```{r}
#| label: winter-prep-data
winter_olympics <- olympics |>
  filter(season == "Winter")
```

## Column - Medals by sport and year {width=65%}

### Row - Medals by sport {height=60%}

```{r}
#| label: winter-medals-by-sport
#| title: Medals by sport
#| fig-width: 10
winter_olympics |>
  mutate(
    sport = fct_lump_n(sport, n = 15),
    sport = fct_infreq(sport),
    sport = fct_rev(sport),
    sport = fct_relevel(sport, "Other", after = 0)
  ) |>
  ggplot(aes(y = sport, fill = medal)) +
  geom_bar() +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = c(0.8, 0.2),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```

### Row - Medals by year {height=40%}

::: {.card title="Medals by year"}
Due to World War II, no olympic games were held in 1940 and 1944.

```{r}
#| label: winter-medals-by-year
#| title: Medals by year
#| fig-asp: 0.3
#| fig-width: 10
winter_olympics |>
  count(year, medal) |>
  ggplot(aes(x = year, y = n, color = medal)) +
   geom_point(size = 0.8) +
  geom_line() +
  guides(color = guide_legend(reverse = TRUE)) +
  scale_x_continuous(breaks = seq(1896, 2020, 8)) +
  labs(
    x = "Year",
    y = NULL,
    color = "Medal"
  ) +
  theme_minimal() +
  theme(
    legend.position = c(0.2, 0.8),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = "white", color = "gray")
  )
```
:::

## Column - Medals by country {width=35%}

### Row - Value boxes {height=30%}

```{r}
#| label: winter-calculate-most-medals
winter_most_golds <- winter_olympics |>
  filter(medal == "Gold") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1)

winter_most_silvers <- winter_olympics |>
  filter(medal == "Silver") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1)

winter_most_bronzes <- winter_olympics |>
  filter(medal == "Bronze") |>
  count(team, sort = TRUE) |>
  slice_head(n = 1)
```

::: {.valuebox icon="award-fill" color="#d4af37"}
Most golds: 

`{r} winter_most_golds$n`

`{r} as.character(winter_most_golds$team)`
:::

::: {.valuebox icon="award-fill" color="#c0c0c0"}
Most silvers: 

`{r} winter_most_silvers$n`

`{r} as.character(winter_most_silvers$team)`
:::

::: {.valuebox icon="award-fill" color="#cd7f32"}
Most bronzes: 

`{r} winter_most_bronzes$n`

`{r} as.character(winter_most_bronzes$team)`
:::

### Row - Tabsets of tables {height=70% .tabset}

::: {.card title="Top 30 total medals"}
Teams sorted in descending order of total medals.

```{r}
#| label: winter-top-30-total-medals
winter_olympics |>
  count(team, medal) |>
  pivot_wider(names_from = medal, values_from = n, values_fill = 0) |>
  mutate(total = Bronze + Gold + Silver) |>
  arrange(desc(total), team) |>
  slice_head(n = 30) |>
  select(Team = team, Gold, Silver, Bronze) |>
  gt() |>
  cols_align(align = "left", columns = Team) |>
  data_color(
    method = "numeric",
    palette = "nord::aurora"
  )
```
:::

::: {.card title="Bottom 30 total medals"}
Teams sorted in ascending order of total medals.
```{r}
#| label: winter-bottom-30-total-medals
winter_olympics |>
  count(team, medal) |>
  pivot_wider(names_from = medal, values_from = n, values_fill = 0) |>
  mutate(total = Bronze + Gold + Silver) |>
  arrange(total, team) |>
  slice_head(n = 30) |>
  select(Team = team, Gold, Silver, Bronze) |>
  gt() |>
  cols_align(align = "left", columns = Team) |>
  data_color(
    method = "numeric",
    palette = "nord::frost"
  )
```
:::
